---
title: "Homework 8"
author: "Bryan Correa Gonzalez"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: spacelab
    keep_md: yes
---

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above.  

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  

## Load the libraries
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(skimr)
library(naniar)
``` 

## Data
**1. For this homework, we will use two different data sets. Please load `amniota` and `amphibio`.**  

`amniota` data:  
Myhrvold N, Baldridge E, Chan B, Sivam D, Freeman DL, Ernest SKM (2015). “An amniote life-history
database to perform comparative analyses with birds, mammals, and reptiles.” _Ecology_, *96*, 3109.
doi: 10.1890/15-0846.1 (URL: https://doi.org/10.1890/15-0846.1).
```{r}
amniota <- read.csv("data/amniota.csv") %>% clean_names()
```

`amphibio` data:  
Oliveira BF, São-Pedro VA, Santos-Barrera G, Penone C, Costa GC (2017). “AmphiBIO, a global database
for amphibian ecological traits.” _Scientific Data_, *4*, 170123. doi: 10.1038/sdata.2017.123 (URL:
https://doi.org/10.1038/sdata.2017.123).
```{r}
amphibio <- read.csv("data/amphibio.csv") %>% clean_names()
```

## Questions  
**2. Do some exploratory analysis of the `amniota` data set. Use the function(s) of your choice. Try to get an idea of how NA's are represented in the data.**  
```{r}
glimpse(amniota)
```

```{r}
summary(amniota)
```


**3. Do some exploratory analysis of the `amphibio` data set. Use the function(s) of your choice. Try to get an idea of how NA's are represented in the data.**  
```{r}
summary(amphibio)
```

```{r}
glimpse(amphibio)
```

**4. How many total NA's are in each data set? Do these values make sense? Are NA's represented by values?**   

The summary stats using naniar states that there are 0 missing values for the amniota data but that is not true. From the data set we can see that there are various -999.00 and -999 values which most likely represent NA's. We will need to fix this.

The summary stats using naniar for the amphibio dataframe suggest more reasonable numbers of missing variables with a total of 170564 NA's. The NA's are represented by NA's.

```{r}
n_miss(amniota)
miss_var_summary(amniota)
```

```{r}
n_miss(amphibio)
miss_var_summary(amphibio)
```

**5. Make any necessary replacements in the data such that all NA's appear as "NA".**   

No changes were needed for the amphibio dataset.

```{r}
amniota <- amniota %>%
  replace_with_na_all(condition = ~.x == -999)
miss_var_summary(amniota)
```

**6. Use the package `naniar` to produce a summary, including percentages, of missing data in each column for the `amniota` data.**  
```{r}
miss_var_summary(amniota)
```

**7. Use the package `naniar` to produce a summary, including percentages, of missing data in each column for the `amphibio` data.**
```{r}
miss_var_summary(amphibio)
```

**8. For the `amniota` data, calculate the number of NAs in the `egg_mass_g` column sorted by taxonomic class; i.e. how many NA's are present in the `egg_mass_g` column in birds, mammals, and reptiles? Does this results make sense biologically? How do these results affect your interpretation of NA's?** 

The results do not make biological sense at for mammals, there is 100% missing but platypi lay eggs. For the reptiles and aves, it make sense that there are atleast some measurements of egg mass. However, aves is missing 50% of the egg mass recording and reptiles are missing 90% which is huge. This makes me question the validity of the data set. 

```{r}
amniota %>% 
  distinct(class)
```

```{r}
amniota %>% 
  select(class, egg_mass_g) %>% 
  group_by(class) %>% 
  miss_var_summary()
```

**9. The `amphibio` data have variables that classify species as fossorial (burrowing), terrestrial, aquatic, or arboreal.Calculate the number of NA's in each of these variables. Do you think that the authors intend us to think that there are NA's in these columns or could they represent something else? Explain.**

I think the NA's represent a 0. They seemed to have made it a binary answer where 1 = the animal lives in that habitat (true) or NA = 0 = the animal does not live in that habitat (false).

```{r}
amphibio_habitat <- amphibio %>% 
  select(fos, ter, aqu, arb) %>% 
  pivot_longer(cols = everything(),
               names_to = "habitat",
               values_to = "true_or_false")
amphibio_habitat %>% 
  group_by(habitat) %>% 
  miss_var_summary()
```

**10. Now that we know how NA's are represented in the `amniota` data, how would you load the data such that the values which represent NA's are automatically converted?**
```{r}
read_csv(file = "data/amniota.csv", na = c("-999")) %>% clean_names()
```

**11. Please complete the class project survey at: [BIS 15L Group Project](https://forms.gle/H2j69Z3ZtbLH3efW6)**

## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences.  