---
title: "Homework 4"
author: "Type Your Name Here"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: spacelab
    keep_md: yes
---

## Instructions
Answer the following questions and/or complete the exercises in RMarkdown. Please embed all of your code and push the final work to your repository. Your report should be organized, clean, and run free from errors. Remember, you must remove the `#` for any included code chunks to run.  

## Load the tidyverse
```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

## Data 
For this assignment, we are going to use built-in data on mammal sleep patterns `msleep`. The data are taken from: V. M. Savage and G. B. West. A quantitative, theoretical framework for understanding mammalian sleep. Proceedings of the National Academy of Sciences, 104 (3):1051-1056, 2007.  

Since the data are built-in, they do not need to be stored as a separate data frame in order to use them.  

**1. Learn about the data and variables by using the help function in R.**
```{r}
help(msleep)
```

**2. What are the names of the variables in msleep?**  
```{r}
names(msleep)
```

**3. Use one of the summary functions you have learned to get an idea of the structure of the data.**  
```{r}
summary(msleep)
```

```{r}
glimpse(msleep)
```

**4. The variable `conservation` categorizes the conservation status of the mammals in the data. How many mammals are endangered?**
It appears like 4 mammals are endangered.
```{r}
msleep
filter(msleep, conservation == "en")
```

**5. Use `filter` to pull out the endangered mammals. Store this as a new object called `endangered`.**
```{r}
endangered <- filter(msleep, conservation == "en")
```

**6. The variable `vore` categorizes the feeding strategy of the mammals in the data. How many mammals are in each category of `vore`?**
There are 83 mammals in total. 19 of which are carnivorous. 32 are herbivorous. 20 are insectivorous. 5 are omnivorous.
```{r}
vore_mammals <- select(msleep, name, vore)
vore_mammals
summary(vore_mammals)
carni_mammals <- filter(vore_mammals, vore == "carni")
herbi_mammals <- filter(vore_mammals, vore == "herbi")
omni_mammals <- filter(vore_mammals, vore == "omni")
insecti_mammals <- filter(vore_mammals, vore == "insecti")
carni_mammals
herbi_mammals
omni_mammals
insecti_mammals
anyNA(vore_mammals)

```
```{r}
table(msleep$vore) # better thing to use
```


**7. Use `filter` to find the insectivore(s) with endangered status.**
```{r}
filter(msleep, vore == "insecti", conservation == "en")
```

**8. We are interested in two groups; small and large mammals. Let's define small as less than or equal to 1kg body weight and large as greater than or equal to 200kg body weight. Make two new dataframes (large and small) based on these parameters.**
```{r}
small_mammals <- filter(msleep, bodywt <= 1)
small_mammals
```
```{r}
large_mammals <- filter(msleep, bodywt>=200)
large_mammals
```


**9. Do large or small animals sleep longer on average?** 
Small mammals sleep longer.
```{r}
mean(small_mammals$sleep_total)
```
```{r}
mean(large_mammals$sleep_total)
```


**10. Which animal sleeps the least among the entire dataframe?**
Giraffes sleep the least.
```{r}
min(msleep$sleep_total)
filter(msleep, sleep_total == 1.9)
```

## Knit and Upload
Please knit your work as a .pdf or .html file and upload to Canvas. Homework is due before the start of the next lab. No late work is accepted. Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  